Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods

### 2017-10-21 thesis/draft.Rmd
conclusion wip

First, we remove randomly remove data such that 20% of the data is missing, then estimate the imputation model. Using this model, we calculate the MARE within the training data.
```{r}
miss_idx <- sample(1:nrow(dfa2), round(nrow(dfa2)*0.2))

@@

  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
               mutate(DayType = ifelse(is_pub_hol,
                                       "Holiday",
                                       daytype(pred_date)),
                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
                      Date = pred_date)

@@

ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),
                            is_pub_hol = FALSE){
  day_df <- data.frame()
      system.time(for(t in 0:23){
        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
      })
  return(as.tibble(day_df))
}

@@

>>>>

First, data is randomly removed such that 20% of the training data is missing, then the imputation model is estimated. Using the estimated model, the MARE within the training data is calculated (in-sample MARE).
```{r}
miss_idx <- sample(1:nrow(dfa2), round(nrow(dfa2)*0.2))

@@

  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
               mutate(DayType = ifelse(is_pub_hol,
                                       "Holiday",
                                       daytype(pred_date)),
                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
                      Date = pred_date,
                      Date_Time = ymd_h(paste(Date, t_hour, sep = " ")))

@@

ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),is_pub_hol = FALSE){
  day_df <- data.frame()
      for(t in 0:23){
        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
      }
  return(as.tibble(day_df))
}

@@

Using these function as a basis, visualisations of pedestrian counts can be easily generated in R:
```{r, echo = TRUE, fig.cap = "Prediction for 21/05/2018 in R using `ped_predict_day()`, then plotting the time series using `ggplot2`. This shows the user friendly nature of the function, where the function returns predictions in a tidy, wide format. No data manipulation is required in order to plot the predicted counts.", fig.height=2}
ped_predict_day("2018-05-21") %>% 
  ggplot() + geom_line(aes(x = Date_Time, y = Southbank))
```

# Conclusion  
Analysis of the Melbourne CBD pedestrian data made publicly available by the City of Melbourne has revealed issues with the dataset with relation to missing values in particular. A major issue which was found when using data sourced from the City of Melbourne's official pedestrian data visualisation (also accessible using `rwalkr::walk_melb()`) is potential for false zero-counts where missing values were given the value of zero. As outliers have major implications on analysis, a correction for the false counts in the data was proposed, by checking for lengths of repeated value sequences.  
An imputation method for the pedestrian data was developed with a theory of sensor locations with a large proportion of missing values require modelling with non-deterministic predictors due to the lack of data. This method involved using neighbouring sensor counts data for prediction. Sensors with a small proportion of missing values used imputation models with only time and date variables as predictors. The proposed algorithm was evaluated in comparison to a single model specification of a generalised linear model with only time and date based variables at each sensor location.
